Theory about Normal Basis on $F(2^k)$

### 2016-12-11 dissertation/abstract.tex
dissertation: committee review version

This dissertation presents new techniques for word-level abstraction
with applications in sequential design verification. By bundling
together $k$ bit-level state-variables into one word-level constraint 
expression, the state-space is construed as solutions (variety) to
a set of polynomial constraints (ideal), modeled over the finite
(Galois) field of $2^k$ elements. Subsequently, techniques from
algebraic geometry -- notable, Gr\"obner basis theory and technology
-- are researched to perform reachability analysis and
verification of sequential circuits. This approach adds a ``word-level
dimension'' to state-space abstraction and verification to make the
process more efficient.

>>>>

This dissertation presents new techniques for word-level abstraction
with applications in sequential design verification. By bundling
together $k$ bit-level state-variables into one word-level constraint 
expression, the state-space is construed as solutions (variety) to
a set of polynomial constraints (ideal), modeled over the finite
(Galois) field of $2^k$ elements. Subsequently, techniques from
algebraic geometry -- notably, Gr\"obner basis theory and technology
-- are researched to perform reachability analysis and
verification of sequential circuits. This approach adds a ``word-level
dimension'' to state-space abstraction and verification to make the
process more efficient.

### 2016-12-11 dissertation/intro.tex
dissertation: committee review version

\section{Hardware Design and Verification Overview}
During the last decades, the level of integration in modern VLSI systems is becoming higher and higher because of the Moore's law.
As a result, an entire system with billions of transistors can be built upon a single chip.
The design process also evolves from manual design with little validation, to 
a formal 3-step procedure which requires collaboration of teams with large number of 
engineers. The 3 major steps are: 1) Design, which is to specify and enter the design intent;
2) Implement, which is to refine the design through various abstraction levels with the assistance of Computer-Aided-Design (CAD)
tools; 3) Verify, which is to verify the correctness of design and implementation.

@@

One way to perform verification is by {\it simulation}. It is defined as the collage of all circuit validation 
methods which apply stimulus on the inputs of circuit model and verify correctness of the outputs.
However, simulation is not a complete solution to circuit verification problems. In modern designs with 
large number of logic components and complicated architectures, it is impractical to simulate all possible 
test vectors. Usually only test vectors that correspond to typical failures are selected in simulation, which 
cannot cover unexpected failure patterns caused by special inputs. The notorious Intel FDIV bug \cite{nicely:FDIV}
is a good example where simulation failed. Failure occurred with input assignments which were rarely used in most divisions. Because of the limitation of simulation,  
test engineers from Intel failed to detect the bug,  which brought about $\$475$ million dollars recalling bill
for the company. Thus,  new methods that can guarantee the correctness of the design are necessary to be explored.

@@

A part of formal verification techniques adopt concepts and algorithms from 
\emph{computer-algebra} and \emph{algebraic geometry}.
Algebraic geometry provides a way to reason about the presence or absence of solutions 
without actually solving the system of constraints.
Using methods in \cite{Avrunin:CAV,condrat-tacas07,gbverify:2007,jinpeng,pruss:tcad15}, 
the circuit design can be transformed into a polynomial system. Subsequently, this system
of polynomials is canonicalized by computing a Gr\"obner basis (GB) \cite{gb_book}. 
Computation of GB allows for 
a straightforward proof of important properties of the polynomial system, 
such as the presence or absence of 
solutions. These properties can also be leveraged for 
verification. The disadvantage of the GB computation method is that its complexity can be doubly 
exponential in the worst case \cite{dube1986complexity}. Thus, directly performing GB computation 
over arbitrary setup is not practical for industry-level applications. However, recent
breakthroughs in computer-algebra hardware verification have shown
that it is possible to overcome the complexity of this computation while
still utilizing the beneficial properties of GB
\cite{lv:phd,tim:phd}.

>>>>

\section{Hardware Design and Verification Overview}
During the past decades, the level of integration in modern VLSI systems is becoming higher and higher because of the Moore's law.
As a result, an entire system with billions of transistors can be built upon a single chip.
The design process also evolves from manual design with little validation, to 
a formal 3-step procedure which requires collaboration of teams with large number of 
engineers. The 3 major steps are: 1) Design, which is to specify and enter the design intent;
2) Implement, which is to refine the design through various abstraction levels with the assistance of Computer-Aided-Design (CAD)
tools; 3) Verify, which is to verify the correctness of design and implementation.

@@

One way to perform verification is by {\it simulation}. It is the collage of all circuit validation 
methods which apply stimulus on the inputs of circuit model and verify correctness of the outputs.
However, simulation is not a complete solution to circuit verification problems. In modern designs with 
large number of logic components and complicated architectures, it is impractical to simulate all possible 
test vectors. Usually only test vectors that correspond to typical failures are selected in simulation, which 
cannot cover unexpected failure patterns caused by special inputs. The notorious Intel FDIV bug \cite{nicely:FDIV}
is a good example where simulation failed. Failure occurred with input assignments which were rarely used in most divisions. Because of the limitation of simulation,  
test engineers from Intel failed to detect the bug,  which brought about $\$475$ million dollars recalling bill
for the company. Thus,  new methods that can guarantee the correctness of the design are necessary to be explored.

@@

Many formal verification techniques adopt concepts and algorithms from 
\emph{computer-algebra} and \emph{algebraic geometry}.
Algebraic geometry provides a way to reason about the presence or absence of solutions 
without actually solving the system of constraints.
Using methods in \cite{Avrunin:CAV,condrat-tacas07,gbverify:2007,jinpeng,pruss:tcad15}, 
the circuit design can be transformed into a polynomial system. Subsequently, this system
of polynomials is canonicalized by computing a Gr\"obner basis (GB) \cite{gb_book}. 
Computation of GB allows for 
a straightforward proof of important properties of the polynomial system, 
such as the presence or absence of 
solutions. These properties can also be leveraged for 
verification. The disadvantage of the GB computation method is that its complexity can be doubly 
exponential in the worst case \cite{dube1986complexity}. Thus, directly performing GB computation 
over an arbitrary setup is not practical for industry-level applications. However, recent
breakthroughs in computer-algebra hardware verification have shown
that it is possible to overcome the complexity of this computation while
still utilizing the beneficial properties of GB
\cite{lv:phd,tim:phd}.

### 2016-12-11 dissertation/normal.tex
dissertation: committee review version

\begin{table}[H]
\label{tab:slimgb}
	\begin{center}
	    \caption{Runtime of Gr\"obner Basis Computation of Agnew's SMPO in Singular using ATO $>$.}\label{tab:ATO}
	    \begin{tabular}{|c|c||c|} 
	        \hline
		Word Size ($k$) & Number of Polynomials ($d$) & Computation Time (seconds)   \\
		\hline
	        $5$	&  $65$  & $20$ \\
		$11$	&  $275$  & $>35,000$  \\
	        \hline
	    \end{tabular}
	\end{center} 
\end{table}

@@

\begin{itemize}
\item First, a set of polynomials is constructed from each logic gates in the circuit, and they generate 
ideal $J_{gates}$. Similarly, ideal $J_{WB}$ is generated by polynomials representing the word-to-bit correspondences.
They are merged as an ideal describing the combinational logic of the circuit: $J_{ckt} = J_{gates} + J_{WB}$.
\item Second, we impose RATO on ideal $J_{ckt}$: all bit-level variables are sorted using a reverse topological order
derived from the circuit structure, followed by word-level NS outputs ({\it e.g.} $R'$ in Figure \ref{fig:sequential}), 
then word-level PS inputs ({\it e.g.} $A,B,R$ in Figure \ref{fig:sequential}).
\item When computing the Gr\"obner basis of $J_{ckt}+J_0$ (adding vanishing polynomials), because of 
RATO, there exists only one pair of polynomials $(f_w,f_g)$ such that $LCM(lm(f_w),lm(f_g)) \neq lm(f_w)\cdot lm(f_g)$.
\item Then we only need to compute the $Spoly$ for $(f_w,f_g)$, and reduce the $Spoly$ by $J_{ckt}+J_0$.
\end{itemize}

@@

Beyond 162-bit fields, our approach is infeasible --
mostly due to the fact that the intermediate 
abstraction polynomial $R$ is very dense and contains high-degree
terms, which can be infeasible to compute. However, it should be noted
that if we do not use the proposed bit-level to word-level
substitution, and compute reduced \Grobner bases with RATO, then our
approach does not scale beyond 33-bit datapaths.

@@

This chapter has described a method to verify sequential Galois field
multipliers over $\Fkk$ using computer algebra and algebraic geometry
based approach. As sequential Galois field circuits perform the
computations over $k$ clock-cycles, verification requires an efficient
approach to unroll the computation, and represent it as a canonical
word-level multi-variate polynomial. Using algebraic geometry, we show
that the unrolling of the computation at word-level can be performed by
\Grobner bases and elimination term orders. Subsequently, we show
that the complex \Grobner basis computation can be eliminated by means
of a bit-level to word-level substitution, which is implemented using
the binomial expansion over Galois fields and Gaussian
elimination. Our approach is able to verify up to 100-bit sequential
circuits, whereas contemporary techniques fail beyond 23-bit
datapaths.

>>>>

\begin{table}[H]
\begin{center}
	    \caption{Runtime of Gr\"obner Basis Computation of Agnew's SMPO in Singular using ATO $>$.}\label{tab:slimgb}
	    \begin{tabular}{|c|c||c|} 
	        \hline
		Word Size ($k$) & Number of Polynomials ($d$) & Computation Time (seconds)   \\
		\hline
	        $5$	&  $65$  & $16$ \\
	        $6$ &  $90$  & $368$ \\
	        $9$ &  $189$ & $43602$ \\
		$11$	&  $275$  & $TO$ ($>48$ hr)  \\
	        \hline
	    \end{tabular}
	\end{center} 
\end{table}

@@

\begin{itemize}
\item First, a set of polynomials is constructed from each logic gate in the circuit, and they generate 
ideal $J_{gates}$. Similarly, ideal $J_{WB}$ is generated by polynomials representing the word-to-bit correspondences.
They are merged as an ideal describing the combinational logic of the circuit: $J_{ckt} = J_{gates} + J_{WB}$.
\item Second, we impose RATO on ideal $J_{ckt}$: all bit-level variables are sorted using a reverse topological order
derived from the circuit structure, followed by word-level NS outputs ({\it e.g.} $R'$ in Figure \ref{fig:sequential}), 
then word-level PS inputs ({\it e.g.} $A,B,R$ in Figure \ref{fig:sequential}).
\item When computing the Gr\"obner basis of $J_{ckt}+J_0$ (adding vanishing polynomials), because of 
RATO, there exists only one pair of polynomials $(f_w,f_g)$ such that $LCM(lm(f_w),lm(f_g)) \neq lm(f_w)\cdot lm(f_g)$.
\item Then we only need to compute the $Spoly$ for $(f_w,f_g)$, and reduce the $Spoly$ by $J_{ckt}+J_0$.
\end{itemize}

@@

Beyond 162-bit fields, our approach is infeasible --
mostly due to the fact that the intermediate 
abstraction polynomial $R$ is very dense and contains high-degree
terms, which can be infeasible to compute. However, it should be noted
that if we do not use the proposed bit-level to word-level
substitution, and compute reduced \Grobner bases with RATO, then our
approach does not scale beyond 11-bit datapaths (Table \ref{tab:slimgb}).

@@

This chapter has described a method to verify sequential Galois field
multipliers over $\Fkk$ using computer algebra and algebraic geometry
based approach. As sequential Galois field circuits perform the
computations over $k$ clock-cycles, verification requires an efficient
approach to unroll the computation, and represent it as a canonical
word-level multi-variate polynomial. Using algebraic geometry, we show
that the unrolling of the computation at word-level can be performed by
\Grobner bases and elimination term orders. Subsequently, we show
that the complex \Grobner basis computation can be eliminated by means
of a bit-level to word-level substitution, which is implemented using
the binomial expansion over Galois fields and Gaussian
elimination. Our approach is able to verify up to 162-bit sequential
circuits, whereas contemporary techniques ({\it e.g.} BDDs) fail beyond 23-bit
datapaths.

### 2016-12-11 dissertation/prelim_GF.tex
dissertation: committee review version

\section{Design a Normal Basis Multiplier on Gate Level}
\label{sec:nbdesign}
The NB multiplier design consists of fewer gates than ordinary StdB multiplier design, even if 
we use the most naive design. However, the modern NB multiplier design has been improved a lot from the 
very first design model proposed by Massey and Omura in 1986 \cite{MasseyOmura}. In order to 
test our approach on practical contemporary circuits, it is necessary to learn the mechanism and design 
routine of several kinds of modern NB multipliers.
\subsection{Sequential Multiplier with Parallel Outputs}
The major benefit of NB multiplier origins from the sequential design. A straightforward design implementing 
the cyclic-shift of operands and $\lambda$-Matrix logic component is the Massey-Omura multiplier.

@@

\subsection{Multiplier not based on $\lambda$-Matrix}
Both Massey-Omura multiplier and Agnew's SMPO rely on the implementation of $\lambda$-Matrix,
which means that they will be identical if unrolled to full combinational circuits. 
After Agnew's work of parallelization, researchers proposed more designs of SMPO, 
some of them jump out of the box and are independent from $\lambda$-Matrix.
One competitive multiplier design of this type is invented by Reyhani-Masoleh and Hasan 
\cite{RHmulti}, which is therefore called RH-SMPO.

@@

\section{Concluding Remarks}
In this chapter, we introduced basic concepts such as the definition of finite fields and the construction 
of finite fields. Moreover, we described a special kind of basis in finite field, and its 
application on Galois field hardware design. The sequential Galois field multipliers based on these 
principles are used as good candidates for applying our functional verification approach in subsequent chapters.

>>>>

\section{Design of a Normal Basis Multiplier on Gate Level}
\label{sec:nbdesign}
The NB multiplier design consists of fewer gates than ordinary StdB multiplier design, even if 
we use the most naive design. However, the modern NB multiplier design has been improved a lot from the 
very first design model proposed by Massey and Omura in 1986 \cite{MasseyOmura}. In order to 
test our approach on practical contemporary circuits, it is necessary to learn the mechanism and design 
routine of several kinds of modern NB multipliers.
\subsection{Sequential Multiplier with Parallel Outputs}
The major benefit of NB multiplier origins from the sequential design. A straightforward design implementing 
the cyclic-shift of operands and $\lambda$-Matrix logic component is the Massey-Omura multiplier.

@@

\subsection{Multiplier not based on the $\lambda$-Matrix}
Both Massey-Omura multiplier and Agnew's SMPO rely on the implementation of $\lambda$-Matrix,
which means that they will be identical if unrolled to full combinational circuits. 
After Agnew's work of parallelization, researchers proposed more designs of SMPO, 
some of them jump out of the box and are independent from $\lambda$-Matrix.
One competitive multiplier design of this type is invented by Reyhani-Masoleh and Hasan 
\cite{RHmulti}, which is therefore called RH-SMPO.

@@

\section{Concluding Remarks}
In this chapter, we introduced basic concepts such as the definition of finite fields and the construction 
of finite fields. Moreover, we described a special kind of basis in finite field, and its 
application on Galois field hardware design. The sequential Galois field multipliers based on these 
principles are used as candidates for applying our functional verification approach in subsequent chapters.

### 2016-12-11 dissertation/prelim_basis.tex
dissertation: committee review version

\section{Algebraic Geometry Fundamentals}
\label{sec:geo}
\subsection{Monomials, Polynomials and Polynomial Arithmetics}
\begin{Definition} \label{def:mono}
A {\bf monomial} in variables $x_1,x_2,\cdots,x_d$ is a product of the form:
\begin{equation}
x_1^{{\alpha}_1} \cdot x_2^{{\alpha}_2} \cdot \cdots x_d^{{\alpha}_d},
\end{equation}
where $\alpha_i \ge 0, i\in\{1,\cdots,d\}$. 
The total degree of the monomial is $\alpha_{1}+\cdots+\alpha_{d}$.
\end{Definition}

@@

The high computational complexity of the Gr\"obner basis problem is an important issue
because of its high cost on time and space. Concretely, for arbitrary polynomial set, 
the worst case computation time/space cost of its Gr\"obner basis is doubly exponential while the total degree of 
polynomials in the Gr\"obner basis is bounded by 
$O(2(\frac{1}{2}d^2+d)^{2^{n-1}})$ where $d$ is the degree of the ideal and $n$ is the number of variables
\cite{dube1986complexity}. However, in practice such as applications of circuit verification, 
the polynomial is well restricted rather than arbitrary.
Gao {\it et al.} \cite{gao:qe-gf-gb} proves that in zero-dimensioned ideals, the complexity reduces to single exponential $q^{O(|\phi|)}$.
This provides the possibility to make use of Gr\"obner basis under restricted situations,
which is the theoretic basis of this dissertation.

@@

\begin{Theorem}[Abstraction Theorem]
Using the notations from Problem Setup \ref{setup:abs} at the beginning of this subsection,
we compute a \Grobner basis $G$ of ideal $(J+J_0)$ using the abstraction term 
order $>$. Then: \\
1) For every word-level input $A_i$, $G$ must contain the vanishing 
polynomial $A_i^q - A_i$ as the only polynomial with $A_i$ as its only 
variable;\\
2) $G$ must contain a polynomial of the form 
$Z + \mathcal{G}(A_1,\dots,A_n)$; and\\ 
3) $Z + \mathcal{G}(A_1,\dots,A_n)$ is such that 
$\Func(A_1,\dots,A_n) = {\mathcal{G}}(A_1,\dots,A_n),
\forall A_1,\dots,A_n \in \Fq$. 
In other words, ${\mathcal{G}}(A_1,\dots,A_n)$ and $\Func(A_1,\dots,A_n)$ are
equal as polynomial functions over $\Fq$.
\label{thm:abs}
\end{Theorem}

>>>>

\section{Algebraic Geometry Fundamentals}
\label{sec:geo}
\subsection{Monomials, Polynomials and Polynomial Algebra}
\begin{Definition} \label{def:mono}
A {\bf monomial} in variables $x_1,x_2,\cdots,x_d$ is a product of the form:
\begin{equation}
x_1^{{\alpha}_1} \cdot x_2^{{\alpha}_2} \cdot \cdots x_d^{{\alpha}_d},
\end{equation}
where $\alpha_i \ge 0, i\in\{1,\cdots,d\}$. 
The total degree of the monomial is $\alpha_{1}+\cdots+\alpha_{d}$.
\end{Definition}

@@

The high computational complexity of the Gr\"obner basis problem is an important issue
because of its high cost on time and space. Concretely, for arbitrary polynomial set, 
the worst case computation time/space cost of its Gr\"obner basis is doubly exponential, as the total degree of 
polynomials in the Gr\"obner basis is bounded by 
$O(2(\frac{1}{2}d^2+d)^{2^{n-1}})$ where $d$ is the degree of the ideal and $n$ is the number of variables
\cite{dube1986complexity}. However, in practice such as applications of circuit verification, 
the polynomial is well restricted rather than arbitrary.
Gao {\it et al.} \cite{gao:qe-gf-gb} proves that in zero-dimensioned ideals, the complexity reduces to single exponential $q^{O(|\phi|)}$.
This provides the possibility to make use of Gr\"obner basis under restricted situations,
which is the theoretical basis of this dissertation.

@@

\begin{Theorem}[Abstraction Theorem]
\cite{tim:phd}
Using the notations from Problem Setup \ref{setup:abs} at the beginning of this subsection,
we compute a \Grobner basis $G$ of ideal $(J+J_0)$ using the abstraction term 
order $>$. Then: \\
1) For every word-level input $A_i$, $G$ must contain the vanishing 
polynomial $A_i^q - A_i$ as the only polynomial with $A_i$ as its only 
variable;\\
2) $G$ must contain a polynomial of the form 
$Z + \mathcal{G}(A_1,\dots,A_n)$; and\\ 
3) $Z + \mathcal{G}(A_1,\dots,A_n)$ is such that 
$\Func(A_1,\dots,A_n) = {\mathcal{G}}(A_1,\dots,A_n),
\forall A_1,\dots,A_n \in \Fq$. 
In other words, ${\mathcal{G}}(A_1,\dots,A_n)$ and $\Func(A_1,\dots,A_n)$ are
equal as polynomial functions over $\Fq$.
\label{thm:abs}
\end{Theorem}

### 2016-12-11 dissertation/prev.tex
dissertation: committee review version

\chapter{Previous Work}
\label{ch:prev}
\section{Sequential Equivalence Checking}
As an important component of formal verification for sequential circuits, SEC techniques 
have been developed over decades and widely utilized in both academia and industry. 
The specification of a sequential circuit can be modeled as a (golden model) state machine,
SEC is performed to compare between the circuit for test and the golden one.
% The n\"aive method to implement SEC is: preload both circuits to the same initial states, 
% and assign their primary inputs to the same values during all clock-cycles.
% This method needs to be operated along with state space traversal,  therefore it is 
% less efficient. Moreover,  most SEC only check the primary outputs/inputs consistency 
% and does not require the 1:1 state correspondence,  so state space traversal is not always necessary.
One way to implement SEC is to create miter with two circuits to be verified, then 
prove that there exists no sequence of inputs that generates different outputs.

Researchers proposed improvements by using Boolean functions to represent 
a set of states/transitions \cite{coudert2003unified, coudert1990verification},  or by dividing the sequential circuit to 
smaller subcircuit and re-model the FSM to conditional FSMs \cite{khasidashvili2004theoretical}. IBM created a toolset with 
interfaces that focuses on only designated initial states and remove redundancies in state space \cite{baumgartner2007scalable}.

@@

Unrolling provides a way to transform a sequential circuit to combinational 
circuit. Therefore,  methods which can be applied to combinational circuit 
verification are also suitable for unrolled sequential circuits. The canonical 
graphical representation of the combinational circuit after unrolling is also 
the canonical representation of the original sequential circuit. For the sequential 
equivalence checking problem, we can also unroll the circuit to be verified and the 
specification to combinational ones, and then perform combinational equivalence checking
techniques \cite{avoj2010combinational}. In the following part we review research and techniques which 
can be applied to unrolled sequential circuits.

@@

The use of DDs in traditional formal verification has a lot of advantages. 
For example, DD-based model checking is very efficient as long as the DDs of sequential 
circuit can be setup. The existence of violating states in constructed DDs 
immediately deduces the violation of property. However, when the design gets
larger and larger, the time and space cost of building and storing the diagram 
increases rapidly. In our experiment of verifying a $k$-bit arithmetic circuit 
using ZDDs, when $k$ is larger than 100, the construction of ZDDs occupies 
over $99\%$ runtime of the whole procedure.
\subsection{Combinational Equivalence Checking Techniques}
The CEC problem can be solved using various methods.
Besides using canonical DDs (BDDs
\cite{BRYA86} and their word-level variants \cite{WLS}),
non-canonical representations such as And-Invert-Graph (AIG) based reductions 
\cite{AIG:2002,alanmi:cec:iccad2006} are also very effective. 
Solvers for satisfiability problems (SAT) are good candidates to solve CEC problems,
as long as the miter of two circuits can be described using conjunctive normal form (CNF)
formulas. Applications of SAT on CEC include circuit-SAT solvers \cite{csat}, {\it etc.}
If the circuits being compared are structurally highly similar, AIG and circuit-SAT 
based approaches are proved to be efficient.
However, when the circuits are functionally equivalent but structurally very dissimilar, none of the 
  contemporary techniques including quantifier-free bit-vector 
  (QF-BV) theory based SMT-solvers \cite{Cryptol:fmcad09},
  offer a practical solution.

Recently integer polynomial based techniques \cite{ciesielski2014function,rolf:date16} are proposed  to verify the functional 
correctness of integer arithmetic circuits. Their approach formulates the output signature as a polynomial function 
with binary variables and integer coefficients, then rewrite the polynomial by substituting gate output with gate inputs
variables. After going through the backward rewriting procedure,  the polynomial
will be composed by only input variables. Then the polynomial is converted to a canonical representation, and compared
with designated input signature. If they are equivalent, then the arithmetic circuit is successfully verified.
This approach incurs of polynomial term explosion during the backward rewriting. The authors proposed a heuristic
to levelize the arithmetic circuit, and substitute several gates' variables at the same time to minimize the risk. However, 
the heuristic is proved to be less effective when the inner symmetry of circuit structure is missing.

@@

$k$-BMC with interpolation is a purely incremental model checking approach, and the interpolation procedure relies
on UNSAT core analysis. To overcome these weaknesses, a hybrid model checker named as IC3 is developed 
\cite{bradley2011sat,bradley2011incremental}. IC3 works incrementally to find inductive subclauses
of negations of reached states, meanwhile it is monolithic when computing over-approximations to sets of reachable
states within $1,2,\dots,k$ steps. It is proved to be more efficient than interpolation based model checking
although using similar mechanisms.

@@

\section{Concluding Remarks}
From the investigation of previous work, techniques are to be researched that 
can operate the FSM traversal on word-level to verify a property excluding spurious 
faults. Meanwhile, many abstraction refinement techniques utilize information from UNSAT cores.
We propose to solve these problems in the context of word-level verification,
with data representation, abstraction and algorithm execution all carried out 
on word-level.

In this dissertation, we propose a purely word-level reachability analysis approach which is never done before.
We achieve it by modeling the transition relations, states and the traversal algorithm at word-level.
We borrow inspirations from \cite{tim:phd,gao:qe-gf-gb} to perform state space abstraction.
Moreover, we find applications of our proposed approach to sequential arithmetic verification which is 
not done before, either.

>>>>

\chapter{Previous Work}
\label{ch:prev}
\section{Sequential Equivalence Checking}
As an important component of formal verification for sequential circuits, SEC techniques 
have been developed over decades and widely utilized in both academia and industry. 
The specification of a sequential circuit can be modeled as a (golden model) state machine,
SEC is performed to compare the functionality between the circuit for test and the golden one.
% The n\"aive method to implement SEC is: preload both circuits to the same initial states, 
% and assign their primary inputs to the same values during all clock-cycles.
% This method needs to be operated along with state space traversal,  therefore it is 
% less efficient. Moreover,  most SEC only check the primary outputs/inputs consistency 
% and does not require the 1:1 state correspondence,  so state space traversal is not always necessary.
One way to implement SEC is to create a miter with two circuits to be verified, then 
prove that there exists no sequence of inputs that generates different outputs.

Researchers proposed improvements by using Boolean functions to represent 
a set of states/transitions \cite{coudert2003unified, coudert1990verification},  or by dividing the sequential circuit to 
into a smaller subcircuit and re-model the FSM to conditional FSMs \cite{khasidashvili2004theoretical}. IBM created a toolset with 
interfaces that focuses on only the designated initial states and remove redundancies in state space \cite{baumgartner2007scalable}.

@@

Unrolling provides a way to transform a sequential circuit to combinational 
circuit. Therefore,  methods which can be applied to combinational circuit 
verification are also suitable for unrolled sequential circuits. The canonical 
graphical representation of the combinational circuit after unrolling is also 
the canonical representation of the original sequential circuit. For the sequential 
equivalence checking problem, we can also unroll the circuit to be verified and the 
specification to combinational ones, and then perform combinational equivalence checking
techniques \cite{savoj2010combinational}. In the following part we review research and techniques which 
can be applied to unrolled sequential circuits.

@@

The use of DDs in traditional formal verification has a lot of advantages. 
For example, DD-based model checking is very efficient as long as the DDs of sequential 
circuit can be setup. The existence of violating states in constructed DDs 
immediately deduces the violation of property. However, when the design gets
larger and larger, the time and space cost of building and storing the diagram 
increases rapidly. In our experiment of verifying a $k$-bit arithmetic circuit 
using ZDDs, when $k$ is larger than 100, the construction of ZDDs occupies 
over $99\%$ runtime of the whole procedure.
\subsection{Combinational Equivalence Checking Techniques}
The CEC problem can be solved using various methods.
Besides using canonical DDs (BDDs
\cite{BRYA86} and their word-level variants \cite{WLS}),
non-canonical representations such as And-Invert-Graph (AIG) based reductions 
\cite{AIG:2002,alanmi:cec:iccad2006} are also very effective. 
Solvers for satisfiability problems (SAT) are good candidates to solve CEC problems,
as long as the miter of two circuits can be described using conjunctive normal form (CNF)
formulas. Applications of SAT on CEC include circuit-SAT solvers \cite{csat}, {\it etc.}
If the circuits being compared are structurally highly similar, AIG and circuit-SAT 
based approaches are known to be efficient.
However, when the circuits are functionally equivalent but structurally very dissimilar, none of the 
  contemporary techniques including quantifier-free bit-vector 
  (QF-BV) theory based SMT-solvers \cite{Cryptol:fmcad09},
  offer a practical solution.

Recently integer polynomial based techniques \cite{ciesielski2014function,rolf:date16} are proposed  to verify the functional 
correctness of integer arithmetic circuits. Their approach formulates the output signature as a polynomial function 
with binary variables and integer coefficients, then rewrites the polynomial by substituting gate output with gate 
inputs. After going through the backward rewriting procedure,  the polynomial
will be composed by only input variables. Then the polynomial is converted to a canonical representation, and compared
with designated input signature. If they are equivalent, then the arithmetic circuit is successfully verified.
This approach incurs polynomial term explosion during the backward rewriting. The authors proposed a heuristic
to levelize the arithmetic circuit, and substitute several gates' variables at the same time to minimize the risk. However, 
the heuristic proved to be less effective when the inner symmetry of circuit structure is missing.

@@

The $k$-BMC with interpolation is a purely incremental model checking approach, and the interpolation procedure relies
on UNSAT core analysis. To overcome these weaknesses, a hybrid model checker named as IC3 is developed 
\cite{bradley2011sat,bradley2011incremental}. IC3 works incrementally to find inductive subclauses
of negations of reached states, meanwhile it is monolithic when computing over-approximations to sets of reachable
states within $1,2,\dots,k$ steps. It is proved to be more efficient than interpolation based model checking
although using similar mechanisms.

@@

\section{Concluding Remarks}
From the investigation of previous work, techniques are to be researched that 
can perform the FSM traversal on word-level to verify a property excluding spurious 
faults. Meanwhile, many abstraction refinement techniques utilize information from UNSAT cores.
We propose to solve these problems in the context of word-level verification,
with data representation, abstraction and algorithm execution all carried out 
on word-level.

In this dissertation, we propose a purely word-level reachability analysis approach which has never been done before.
We achieve this by modeling the transition relations, states and the traversal algorithm at word-level.
We borrow inspirations from \cite{tim:phd,gao:qe-gf-gb} to perform state space abstraction.
Moreover, we demonstrate applications of our proposed approach to sequential arithmetic verification which is 
not done before, either.
Finally, we show algebraic geometry analogs of UNSAT cores of polynomial ideals, and describe algorithms
to extract and refine these cores.
